---
title: "Week 3 Assignment Data 605"
author: "Tyler Baker"
date: "2/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

# Week 3 Assignment

## Problem set 1

1) What is the rank of matrix A?

By turning matrix A into rref, one can simply count the number of pivots to find the rank.
```{r rank of A}
a <- matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3), nrow=4,ncol=4, byrow=FALSE)
rref_a <- rref(a)
print(rref_a)
```
Thus, the rank of matrix A is 4.

2) Given an mxn matrix where m>n, what can be the maximum rank? The minimum rank, assuming the matrix is non-zero?

The rank of A = the rank of A transpose.
Let's examine a matrix where the amount of columns is larger than the amount of rows.
```{r}
example_mat <- matrix(c(1,0,0,0,0,1,0,0,0,0,1,0,0,0,0,1,0,0,0,0), nrow=4, ncol=5, byrow=FALSE)
print(example_mat)
```
Here, example_mat is already in rref. We clearly see that there are only 4 pivots. Thus, we know that its rank is equal to 4. Thus, in general we say that given an mxn matrix, the maximum possible rank is equal to the smaller value of m and n.
The minimum rank of a non-zero matrix is 1. The rank of a matrix is the count of linearly independent vectors of a matrix. A non-zero matrix must carry some information to at the very least give us a line. Thus, the minimum rank is again 1.

3)What is the rank of B?
```{r}
b<- matrix(c(1,3,2,2,6,4,1,3,2), nrow=3, ncol=3, byrow=FALSE)
rref_b <- rref(b)
print(rref_b)
```
This shows that the rank of B is 1. Similarly, we can get the same answer by studying the rows of matrix B. We see that row2 is equal to row1 x 3. 
Similarly, row3 is equal to row1 x 2. Thus, row2 and row3 are linear dependent on row1. The same method can similarly be done by examining the columns of B. 

## Problem set 2

The solution to problem set 2 can be seen at the link below.

https://github.com/tylerbaker01/Data-605-Week-3-Assignment/blob/main/week3ps2.png

